---
title: "Feature Mean Time Trend Viz"
author: "Jiaxing Joy Qiu"
date: "4/24/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Documents/BSI/code/Transplant")
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=FALSE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.dim = c(10,3),     # set figure width
                      out.width = "100%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE,
                      result='asis')     # do not show R messages

```

```{r}
# --- prepare suitable dataset for this task  ----
source("./prepare/prep_raw.R")
# --- prepare local functions if in need ----
source("./task_time_trend_inference/functions.R")
```

### Description

* To characterise the difference of a feature over time in 4 primary groups, non_neg / noo_pos / txp_neg / txp_pos, we use the **mean** statistic of the feature at a certain time.

* Visualization comparison
  + Left: empirical
    - the error bar is 95% confidence interval of feature **mean** at given time, calculated from **standard error (se)**, (note: larger the sample size *n*, smaller the *se*)
  + Right: linear regression 
    - the ribbon is 95% confidence interval of estimated **mean response**

```{r}
#  prepare global constants
features <- c( 'Temp', 'Resp', 'Pulse', 'SBP', 'DBP', 'SpO2',
                'GCS', 'FiO2', 'O2_Flow', 'GLUCOSE', 'TOTAL_BILIRUBIN', 'POTASSIUM',
                'ALBUMIN', 'CALCIUM', 'SODIUM', 'WHITE_BLOOD_CELL_COUNT', 'PHOSPHORUS',
                'PROTIME_INR', 'CREATININE', 'PLATELET_COUNT', 'ALT_GPT', 'CO2',
                'ALKALINE_PHOSPHATASE', 'AST_GOT', 'PCO2', 'CHLORIDE', 'TROPONIN_I',
                'PARTIAL_THROMBOPLASTIN_TIME', 'LACTIC_ACID', 'BLOOD_UREA_NITROGEN',
                'OXYGEN_SATURATION', 'MAGNESIUM', 'FIO2')


df_mdl <- bind_rows(df_ep_raw, df_safe_window_raw)
df_mdl$group <- as.factor(df_mdl$group)
df_mdl$cluster <- as.factor(paste0(df_mdl$id,"_",df_mdl$relative_time)) # repeated measures caused by multiple episodes from same patient
df_mdl$relative_hour <- df_mdl$relative_time/4


dd <- datadist(df_mdl)
options(datadist='dd')
options(na.action = "na.omit")
```

```{r}

for (fea in features){
  
  # empirical  time trend plot
  summ <- summarySE(df_mdl, measurevar=fea, groupvars=c("group", "relative_hour"), na.rm = TRUE)
  pd <- position_dodge(0.1) # move them .05 to the left and right
  p1 <- ggplot(summ, aes(x=relative_hour, y=summ[,fea], colour=group)) + 
    geom_errorbar(aes(ymin=summ[,fea]-se, ymax=summ[,fea]+se), size=0.3, width=.05, alpha=0.5, position=pd) +
    geom_line(position=pd) +
    geom_point(size=0.3,position=pd) +
    ylab(fea)+
    scale_color_brewer(palette="Set2")


  # linear regression time trend
  fml <- formula(paste0(fea," ~ rcs(relative_hour,5) * catg(group)"))
  mdl <- robcov(ols(fml,x=TRUE, y=TRUE,data=df_mdl), cluster=df_mdl$cluster)
  p2 <- ggplot(Predict(mdl, relative_hour, group))+
    scale_color_brewer(palette="Set2")

  print(ggpubr::ggarrange(p1,p2, ncol=2, nrow=1,
                          legend="right", common.legend=TRUE))
  
}
```


