---
title: "Longitudinal Analysis (Model-based approach)"
subtitle: "Generalized Least Squares"
author: "Jiaxing Joy Qiu"
date: "4/28/2021"
output: html_document
---
***
### Description

#### Tech
* Regression modeling strategy
  + we use a property to characterise a response within groups
  + after specifying a regression structure, we optimize the model by some criteria 
    - AIC, BIC ... for time-consuming models + large-scaled data
    - AUROC, AUPRC ... for prediction-focused + smaller-size data
  + if assumptions are satisfied with the optimized model, we are allowed to do inference on estimated responce and effects.

* Longitudinal Data (serial data/repeated measures) Analysis
  + Model-based 
    - General Least Squares (GLS)
    - Mixed Effect Model
    - Bayesian hierarchical models
  + Data-based
    - Summary Measures
 
#### Medical
* 4 primary groups
  + non_pos: postive episodes from no-transplant patients
  + non_neg: test-negative and baseline time windows from 'pure negative' no-transplant patients
  + txp_pos: postive episodes from transplant patients
  + txp_neg: test-negative and baseline  time windows from 'pure negative' transplant patients

* An *episode*
  + from 24 hours before to 24 hours after a blood stream acquisition

* Questions 
  1. What's the response trend of a feature during one 'episode' **within** groups?
  2. What's the difference of response trend of a feature during one 'episode' **across** groups?
    

***
### Report
* Assumption-satisfied Variables
  + Temp
  + Resp
  + Pulse
  + SBP
  + DBP
  + POTASSIUM
  + ALBUMIN
  + CALCIUM
  + SODIUM
  + WHITE_BLOOD_CELL_COUNT
  + PHOSPHORUS
  + PLATELET_COUNT
  + CO2
  + PCO2
  + CHLORIDE
  + BLOOD_UREA_NITROGEN
  + MAGNESIUM

* model assumption checking plots
* partial effect plots
* contrast (subtract) inference plots
  + by subtracting estimated response from 2 groups and comparing the confidence interval with 0, we can do inference (example - Temperature)
    - In no-transplant group, our data suggest a significant increase in temperature in postive episodes than control (negative+baseline) episodes
    - In transplant group, our data **don't** suggest a significant increase in positive episodes than control (negative-baseline) episodes

***
```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_knit$set(root.dir = "~/Documents/BSI/code/Transplant")
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=FALSE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.dim = c(10,7),     # set figure width
                      out.width = "100%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE,
                      results='asis')     # do not show R messages

```

```{r}
# --- prepare suitable dataset for this task  ----
source("./prepare/prep_raw.R")
# --- prepare global functions ----
source("./prepare/functions.R")
```

```{r}
# --- prepare local functions and packages ----
require(rms)
require(nlme)
require(ggpubr)

# ---- prepare feature names ----
features <- c('Temp', 
              'Resp', 
              'Pulse', 
              'SBP', 
              'DBP', 
              #'SpO2',
              #'GCS', 
              #'FiO2', 
              #'O2_Flow', 
              #'GLUCOSE', 
              #'TOTAL_BILIRUBIN', 
              'POTASSIUM',
              'ALBUMIN', 
              'CALCIUM', 
              'SODIUM', 
              'WHITE_BLOOD_CELL_COUNT', 
              'PHOSPHORUS',
              #'PROTIME_INR', 
              #'CREATININE', 
              'PLATELET_COUNT', 
              #'ALT_GPT', 
              'CO2',
              #'ALKALINE_PHOSPHATASE', 
              #'AST_GOT', 
              'PCO2', 
              'CHLORIDE', 
              #'TROPONIN_I',
              #'PARTIAL_THROMBOPLASTIN_TIME', 
              #'LACTIC_ACID', 
              'BLOOD_UREA_NITROGEN',
              #'OXYGEN_SATURATION', 
              'MAGNESIUM'#, 
              #'FIO2'
              )
# ---- prepare task data ----
data <- bind_rows(df_ep_raw, df_safe_window_raw)
data$group <- as.factor(data$group)
data$ep_id <- as.factor(data$ep_id)
data$time_idx <- as.numeric(data$relative_time - min(data$relative_time) +1) # start 1 step 1 time index ##### don't name variable 'time'
```


```{r}

for (fea in features){
  # ---- print html title -----
  cat("\n***\n")
  cat("\n####", fea ,"mean response\n") 

  # ---- prepare model object arguments ----
  df_mdl <- data[, c(fea, 'time_idx', 'group', 'ep_id')]
  df_mdl$y <- df_mdl[, fea]
  dd <- datadist(df_mdl)
  options(datadistt='dd')
  options(prType='html')
  fml <- formula("y ~ rcs(time_idx, 5) * catg(group)")
  
  
  # ---- load pretrained model object ----
  load(file = paste0("./task_longitudinal_mdl/gls/gls_mdl_obj/",fea,"_mdl.rda"))
  
  # ---- model assumption plots ----
  cat("\n* Check model assumptions\n") 
  
  df_res <- na.omit(df_mdl)
  df_res$resid <- r<- resid(mdl)
  df_res$fitted <- fitted(mdl)
 
  p1 <- ggplot(df_res, aes(x=fitted, y=resid)) + geom_point(size=0.1) + 
    facet_grid(~ group) + ylab('Residuals')
  
  p3 <- ggplot(df_res, aes(sample=resid)) + stat_qq(size=0.1) + geom_abline(intercept = mean(r), slope = sd(r)) + ylab('Residuals')
  p4 <- ggplot(df_res, aes(x=(as.numeric(time_idx)-97)/4, y=resid)) +
    #stat_summary(fun.data="mean_sdl", geom = "smooth") + 
    geom_point(size=0.1) +
    ylab('Residuals') + xlab('relative hour')
  p2 <- ggpubr::ggarrange(p3, p4, nrow=1, ncol=2)
  print(ggpubr::ggarrange(p1, p2, nrow=2, ncol=1))
  
  
  
  # ----  partial effect and contrast inference plots ----
  cat("\n* Partiel effect and contrast inference\n")
  p_eff <- ggplot(Predict(mdl, time_idx, group)) + ylab(fea) 
  #p <- Predict(mdl, time_idx, group)
  #s <- Predict(mdl, time_idx, group, conf.type = 'simultaneous')
  #ps <- rbind(Pointwise=p, Simultaneous=s)
  
  # ---- contrast dataframes ----
  # non_pos - non_neg
  t <- seq(1, 197)
  k <- contrast(mdl,
                list(time_idx=t, group = 'non_pos'),
                list(time_idx=t, group = 'non_neg'))
  k_non <- as.data.frame(k[c('time_idx', 'Contrast', 'Lower', 'Upper')])
  
  # txp_pos - txp_neg
  t <- seq(1, 197)
  k <- contrast(mdl,
                a=list(time_idx=t, group = 'txp_pos'),
                b=list(time_idx=t, group = 'txp_neg'))
  k_txp <- as.data.frame(k[c('time_idx', 'Contrast', 'Lower', 'Upper')])
 
  # txp_pos - non_pos
  t <- seq(1, 197)
  k <- contrast(mdl,
                list(time_idx=t, group = 'non_pos'),
                list(time_idx=t, group = 'txp_pos'))
  k_pos <- as.data.frame(k[c('time_idx', 'Contrast', 'Lower', 'Upper')])
  
  # txp_neg - non_neg 
  t <- seq(1, 197)
  k <- contrast(mdl,
                list(time_idx=t, group = 'non_neg'),
                list(time_idx=t, group = 'txp_neg'))
  k_neg <- as.data.frame(k[c('time_idx', 'Contrast', 'Lower', 'Upper')])
  
  
  
  # ---- contrast plots ----
  # coerce same y lim
  l <- min(k_non$Lower, k_txp$Lower, k_pos$Lower, k_neg$Lower)
  u <- max(k_non$Upper, k_txp$Upper, k_pos$Upper, k_neg$Upper)
  yl <- ylim(l,u)
  
  p_non <- ggplot(k_non, aes(x=(time_idx-97)/4, y=Contrast)) + 
    geom_line() +
    geom_ribbon(aes(ymin=Lower, ymax=Upper), alpha=0.2, linetype=0, show_guide=FALSE) + 
    geom_hline(yintercept = 0, col='red', size=0.2)+
    xlab("relative hour") +
    ylab("pos - neg") +
    yl +
    ggtitle("No-transplant")
  
  p_txp <- ggplot(k_txp, aes(x=(time_idx-97)/4, y=Contrast)) + 
    geom_line() +
    geom_ribbon(aes(ymin=Lower, ymax=Upper), alpha=0.2, linetype=0, show_guide=FALSE) + 
    geom_hline(yintercept = 0, col='red', size=0.2)+
    xlab("relative hour") +
    ylab("pos - neg") +
    yl +
    ggtitle("Transplant")
  
  p_pos <- ggplot(k_pos, aes(x=(time_idx-97)/4, y=Contrast)) + 
    geom_line() +
    geom_ribbon(aes(ymin=Lower, ymax=Upper), alpha=0.2, linetype=0, show_guide=FALSE) + 
    geom_hline(yintercept = 0, col='red', size=0.2)+
    xlab("relative hour") +
    ylab("non - txp") +
    yl +
    ggtitle("Positive")
  
  p_neg <- ggplot(k_neg, aes(x=(time_idx-97)/4, y=Contrast)) + 
    geom_line() +
    geom_ribbon(aes(ymin=Lower, ymax=Upper), alpha=0.2, linetype=0, show_guide=FALSE) + 
    geom_hline(yintercept = 0, col='red', size=0.2)+
    xlab("relative hour") +
    ylab("non - txp") +
    yl +
    ggtitle("Negative")
  
  p_diff <- ggpubr::ggarrange(p_non, p_txp, p_pos, p_neg, nrow=2, ncol=2)
  print(ggpubr::ggarrange(p_eff, p_diff, nrow=2, ncol=1))

}

```



