---
title: "OLR Estimated Mean Trend (Modeling and Assumption check)"
author: "Jiaxing Joy Qiu"
date: "4/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Documents/BSI/code/Transplant")
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=FALSE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.dim = c(10,6),     # set figure width
                      out.width = "100%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE,
                      result='asis')     # do not show R messages

```

```{r}
# --- prepare suitable dataset for this task  ----
source("./prepare/prep_raw.R")
# --- prepare global functions ----
source("./prepare/functions.R")
# --- prepare local functions if in need ----
```

## Description

### Hypothesis

* Mean of physiological and clinical features change **differently** over time in 4 primary groups - txp_pos, non_pos, txp_neg, non_neg. Our goal is to quantify the difference statistically.

### Interpretation

* **If linear regression assumptions holds**, the departure of confidence interval of feature mean conditioned/given a certain time in 4 groups reviews "significant" difference in time trend of feature mean in 4 groups. 

### Data

* time window
  + (-24h,24h) around a blood culture
  + 48h window around an observation with the most number of data available in a patient's safe zone

* Imputation
  + raw

### Method

* Ordinary linear regression
  + Formula 
    - feature ~ rcs(time, 5) * group
  + Cluster
    - multiple episodes from one patient

    
  + Outcome(y-axis)
    - the estimated conditional feature mean value with 95% confidence interval for the estimation
  
  + Assumptions checking
    - **L**inearity: responce feature is a linear combination of predictors (r vs. fitted)
    - **I**ndependence: the errors are independent (conditioned on time) (r vs. fitted)
    - **N**ormality: the distribution of the errors should follow a normal distribution (qqplot)
    - **E**qual Variance: the error variance is the same at any set of predictors (r vs. predictors)
  


```{r}

#  prepare global constants
features <- c( 'Temp', 'Resp', 'Pulse', 'SBP', 'DBP', 'SpO2',
                'GCS', 'FiO2', 'O2_Flow', 'GLUCOSE', 'TOTAL_BILIRUBIN', 'POTASSIUM',
                'ALBUMIN', 'CALCIUM', 'SODIUM', 'WHITE_BLOOD_CELL_COUNT', 'PHOSPHORUS',
                'PROTIME_INR', 'CREATININE', 'PLATELET_COUNT', 'ALT_GPT', 'CO2',
                'ALKALINE_PHOSPHATASE', 'AST_GOT', 'PCO2', 'CHLORIDE', 'TROPONIN_I',
                'PARTIAL_THROMBOPLASTIN_TIME', 'LACTIC_ACID', 'BLOOD_UREA_NITROGEN',
                'OXYGEN_SATURATION', 'MAGNESIUM', 'FIO2')

```

```{r}
# prepare modeling data
df_mdl <- bind_rows(df_ep_raw, df_safe_window_raw)
df_mdl$group <- as.factor(df_mdl$group)
df_mdl$cluster <- as.factor(paste0(df_mdl$id,"_",df_mdl$relative_time)) # repeated measures caused by multiple episodes from same patient
df_mdl$relative_hour <- df_mdl$relative_time/4


dd <- datadist(df_mdl)
options(datadist='dd')
options(na.action = "na.omit")
```


```{r results='asis'}
for (fea in features){
  fml <- formula(paste0(fea," ~ rcs(relative_hour,5) *catg(group)"))
  mdl <- robcov(ols(fml,x=TRUE, y=TRUE,data=df_mdl), cluster=df_mdl$cluster)
  print(ggplot(Predict(mdl, relative_hour, group)))
  
  # check normality of residuals
  r <- resid(mdl)
  par(mfrow=c(2,2))
  plot(fitted(mdl), r); abline(h=0)  # check independence of errors and linearity
  with(df_mdl, plot(relative_hour, r)); abline(h=0) # check equal variances
  with(df_mdl, plot(group, r)); abline(h=0) # check equal variances
  qqnorm(r) # check normality
  qqline(as.numeric(r))
  par(mfrow=c(1,1))
}
```
