---
title: "OLR Estimated Mean Trend (Inference)"
author: "Jiaxing Joy Qiu"
date: "4/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Documents/BSI/code/Transplant")
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=FALSE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.dim = c(10,6),     # set figure width
                      out.width = "100%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE,
                      result='asis')     # do not show R messages

```

```{r}
# --- prepare suitable dataset for this task  ----
source("./prepare/prep_org.R")
# --- prepare global functions ----
source("./prepare/functions.R")
# --- prepare local functions if in need ----
```

### Regression Modeling Strategy

* To charaterise clinical and physiological features within a cohort of interest, we use a property to represent them.
  + In OLR (oridinary linear regression), the property is mean
  + In QR (quantile regression), the property is median or quantiles
  + In LR (logistic regression), the property is the probability of a binary outcome = 1

* If assumptions are satisfied in above modeling strategies
  + we are allowed to do inference / test hypotheses using the model results


### OLR assumption-statisfied features

+ Temp
+ Pulse
+ SBP
+ DBP
+ POTASSIUM
+ ALBUMIN
+ CALCIUM
+ WHITE_BLOOD_CELL_COUNT
+ CO2
+ CHLORIDE
+ MAGNESIUM


### Hypotheses to test (Inference)
1. Clinical and physiological features **change** over time in each group - txp_pos / non_pos / txp_neg / non_neg



### Data

* time window
  + (-24h,24h) around a blood culture
  + 48h window around an observation in a patient's safe zone which has most number of data available

* Imputation
  + individual 




```{r}
# prepare modeling data
df_mdl <- bind_rows(df_ep_original, df_safe_window_original)
df_mdl$group <- as.factor(df_mdl$group)
df_mdl$cluster <- as.factor(paste0(df_mdl$id,"_",df_mdl$relative_time)) # repeated measures caused by multiple episodes from same patient
df_mdl$relative_hour <- df_mdl$relative_time/4

```

```{r}
#  prepare features that satisfy OLS 
features <- c( 'Temp','Pulse', 'SBP', 'DBP', 'POTASSIUM',
                'ALBUMIN', 'CALCIUM', 'WHITE_BLOOD_CELL_COUNT', 'CO2', 'CHLORIDE','MAGNESIUM')

```

```{r results='asis'}
dd <- datadist(df_mdl)
options(datadist='dd')
options(na.action = "na.omit")

for (fea in features){
  fml <- formula(paste0(fea," ~ rcs(relative_hour,5) *catg(group)"))
  mdl <- robcov(ols(fml,x=TRUE, y=TRUE,data=df_mdl), cluster=df_mdl$cluster)
  print(ggplot(Predict(mdl, relative_hour, group)))
  
  par(mfrow=c(1,3))
  plot(mdl$fitted.values, mdl$residuals)
  qqnorm(mdl$residuals)
  qqline(mdl$residuals)
  hist(mdl$residuals)
  par(mfrow=c(1,1))
}
```
